---
output: github_document
---

# paradox

[![CRAN](https://www.r-pkg.org/badges/version/paradox)](https://cran.r-project.org/package=paradox)
[![Build Status Linux](https://travis-ci.org/mlr-org/paradox.svg?branch=master)](https://travis-ci.org/mlr-org/paradox)
[![Build Status Windows](https://ci.appveyor.com/api/projects/status/m26qhpq99cka8l1b?svg=true)](https://ci.appveyor.com/project/jakob-r/paradox)
[![Coverage Status](https://coveralls.io/repos/github/mlr-org/paradox/badge.svg?branch=master)](https://coveralls.io/github/mlr-org/paradox?branch=master)


```{r setup, include=FALSE}
set.seed(123)
knitr::opts_chunk$set(cache = FALSE, collapse = TRUE, comment = "#>")
library(paradox)
```

Universal Parameter Space Description and Tools

* [Introduction](https://mlr-org.github.io/paradox/articles/paradox.html)
* [Issues and Bugs](https://github.com/mlr-org/paradox/issues)
* [Documentation](https://mlr-org.github.io/paradox/)

**For an exhaustive introduction, please take a look at the [mlr3book](https://mlr3book.mlr-org.com).**

## Installation
```{r inst, eval=FALSE}
remotes::install_github("mlr-org/paradox", dependencies = TRUE)
```

## Usage

Create a simple ParamSet using all supported Parameter Types:

* integer numbers ("`Int`")
* real-valued numbers ("`Dbl`")
* truth values `TRUE` or `FALSE` ("`Lgl`")
* categorical values from a set of possible strings ("`Fct`")
* further types are only possible by using transformations.

```{r ps}
ps = ParamSet$new(
  params = list(
    ParamInt$new(id = "z", lower = 1, upper = 3),
    ParamDbl$new(id = "x", lower = -10, upper = 10),
    ParamLgl$new(id = "flag"),
    ParamFct$new(id = "methods", levels = c("a","b","c"))
  )
)
```

Draw random samples / create random design:

```{r pssample}
generate_design_random(ps, 3)
```

Generate LHS Design:

```{r pslhs}
generate_design_lhs(ps, 3)
```

Generate Grid Design:

```{r psgrid, R.options=list(max.print=30)}
generate_design_grid(ps, resolution = 2)
```

Properties of the parameters within the ParamSet:

```{r psprobs}
ps$ids()
ps$levels
ps$nlevels
ps$is_number
ps$lower
ps$upper
```

### Parameter Checks

Check that a parameter satisfies all conditions of a `ParamSet`, using `$test()` (returns `FALSE` on mismatch), `$check()` (returns error description on mismatch), and `$assert()` (throws error on mismatch):

```{r, error = TRUE}
ps$test(list(z = 1, x = 1))
ps$test(list(z = -1, x = 1))
ps$check(list(z = -1, x = 1))
ps$assert(list(z = -1, x = 1))
```

### Transformations

Transformations are functions with a fixed signature.

* `x` A named list of parameter values
* `param_set` the `ParamSet` used to create the design

Transformations can be used to change the distributions of sampled parameters. For example, to sample values between 2^-3 and 2^3 in a log2-uniform distribution, one can sample uniformly between -3 and 3 and exponentiate the random value inside the transformation.

```{r pstransscale}
ps = ParamSet$new(
  params = list(
    ParamInt$new(id = "z", lower = -3, upper = 3),
    ParamDbl$new(id = "x", lower = 0, upper = 1)
  )
)
ps$trafo = function(x, param_set) {
  x$z = 2^x$z
  return(x)
}
ps_smplr = SamplerUnif$new(ps)
x = ps_smplr$sample(2)
xst = x$transpose()
xst
```
