---
output: github_document
---
# phng

[![Build Status Linux](https://travis-ci.org/mlr-org/phng.svg?branch=master)](https://travis-ci.org/mlr-org/phng)
[![Build Status Windows](https://ci.appveyor.com/api/projects/status/m26qhpq99cka8l1b?svg=true)](https://ci.appveyor.com/project/jakob-r/phng)
[![Coverage Status](https://coveralls.io/repos/github/mlr-org/phng/badge.svg?branch=master)](https://coveralls.io/github/mlr-org/phng?branch=master)


```{r setup, include=FALSE}
set.seed(123)
knitr::opts_chunk$set(cache = TRUE, collapse = FALSE, warning = FALSE, error = FALSE)
library(phng)
```

Universal Parameter Space Description and Tools

* [Issues and Bugs](https://github.com/mlr-org/phng/issues)
* [Documentation](https://mlr-org.github.io/phng)

## Installation
```{r inst, eval=FALSE}
devtools::install_github("mlr-org/phng", dependencies = TRUE)
```

## Usage

Create a simple ParamSet using all supported Parameter Types:

* _Int_egers
* _Real_-valued numbers
* _Flag_ for `TRUE`/`FALSE`
* _Categorical_ values, namely characters.
* Further types are only possible by using transformations.

```{r ps}
ps = ParamSetFlat$new(
  params = list(
    ParamInt$new(id = "z", lower = 1, upper = 3),
    ParamReal$new(id = "x", lower = -10, upper = 10),
    ParamFlag$new(id = "switch"),
    ParamCategorical$new(id = "methods", values = c("a","b","c"))
  )
)
```

Draw random samples / create random design:
```{r pssample}
ps$sample(3)
```

Generate LHS Design:
```{r pslhs}
ps$generateLHSDesign(3)
```

Generate Grid Design:
```{r psgrid}
ps$generateGridDesign(resolution = 2)
```

Properties of the parameters within the ParamSet:
```{r psprobs}
ps$values
ps$param.classes
ps$nlevels
```

### Numeric ParamSet

Things you can do on an all numeric ParamSet:
```{r psnumeric}
ps = ParamSetFlat$new(
  params = c(
    list(ParamInt$new(id = "z", lower = -10, upper = 10)),
    repeatParam(2, ParamReal$new(id = "x", lower = 0, upper = 1))
  )
)

ps$lower
ps$upper
ps$range
```

The usage of `repeatParam` generates tags that indicate to which group the parameters belong to:

```{r psnumerictags}
ps$member.tags
```

This becomes useful if you want to do operations on parameters of one group like with transformations.

### Transformations

Transformations are functions with a fixed signature.

* `x` A `data.table` of parameter values. Each column contains a vector of parameter values of a single parameter.
* `dict` An environment, that can be accessed using the `$` operator. It can contains values that don't belong to any parameter but are important for transformations.
* `tags` A list of the tags for each parameter. Each parameter can have various tags indicating additional characteristics.

Transformations are useful to scale parameters:

```{r pstransscale}
ps = ParamSetFlat$new(
  params = list(
    ParamInt$new(id = "z", lower = -3, upper = 3),
    ParamReal$new(id = "x", lower = 0, upper = 1)
  ),
  trafo = function(x, dict, tags) {
    x$z = 2^x$z
    x$x = round(x$x * dict$p)
    return(x)
  }
)
(x = ps$sample(3))
```

The transformation uses the dictionary and will fail if none is supplied:

```{r pstransscale2, error = TRUE}
ps$transform(x)
```

The dictionary can always be changed:
```{r pstransscale3}
ps$dictionary = list(p = 10)
ps$transform(x)
ps$dictionary = list(p = 1000)
ps$transform(x)
```
### Advanced Transformations

The following creates a ParamSet with a transformation that scales the `x` values and returns them as a vector.
The original parameters will be removed from the trafo result.
Keep in mind that `z` stays untouched and remains after the transformation.

```{r psnumtrans}
ps = ParamSetFlat$new(
  params = c(
    list(ParamInt$new(id = "z", lower = -10, upper = 10)),
    repeatParam(2, ParamReal$new(id = "x", lower = 0, upper = 1))
  ),
  trafo = function(x, dict, tags) {
    scale = function(x1, x2) c(x1, x2) / sqrt(x1^2+x2^2) 
    x$x = Map(scale, x$x.repeated.1, x$x.repeated.2)
    x$x.repeated.1 = NULL
    x$x.repeated.2 = NULL
    return(x)
  }
)
```

The output of all value generating functions won't change for a ParamSet that has a `trafo` function.
Instead these outputs can be put into `ps$transform()` to obtain the desired parameter values.

```{r psnumtrans2}
x = ps$generateLHSDesign(3)
ps$transform(x)
```

For more advanced transformations on repeated parameters you can use `trafoOnRepeatedParam()`:

```{r trafoOnRepeatedParam}
ps = ParamSetFlat$new(
  params = c(
    list(
      ParamFlag$new(id = "switch"),
      ParamInt$new(id = "z", lower = 1, upper = 4)),
    repeatParam(4, ParamReal$new(id = "x", lower = 0, upper = 1))
  ),
  trafo = trafoOnRepeatedParam(
    fun = function(x, dict, tags) {
      scale = function(z, ...) {
        x = c(...)[1:z]
        x / sum(x)
      }
      res = do.call(Map, c(list(f = scale, z = dict$z), as.list(x)))
      list(x = res)
    }, repeated.param.id = "x", additional.params = "z")
  )
(x = ps$sample(3))
ps$transform(x)
```
