---
output: github_document
---
# paradox

[![Build Status Linux](https://travis-ci.org/mlr-org/paradox.svg?branch=master)](https://travis-ci.org/mlr-org/paradox)
[![Build Status Windows](https://ci.appveyor.com/api/projects/status/m26qhpq99cka8l1b?svg=true)](https://ci.appveyor.com/project/jakob-r/paradox)
[![Coverage Status](https://coveralls.io/repos/github/mlr-org/paradox/badge.svg?branch=master)](https://coveralls.io/github/mlr-org/paradox?branch=master)


```{r setup, include=FALSE}
set.seed(123)
knitr::opts_chunk$set(cache = TRUE, collapse = FALSE, warning = FALSE, error = FALSE)
library(paradox)
```

Universal Parameter Space Description and Tools

* [Issues and Bugs](https://github.com/mlr-org/paradox/issues)
* [Documentation](https://mlr-org.github.io/paradox)

## Installation
```{r inst, eval=FALSE}
devtools::install_github("mlr-org/paradox", dependencies = TRUE)
```

## Usage

Create a simple ParamSet using all supported Parameter Types:

* _int_egers
* _real_-valued numbers
* _flag_ for `TRUE`/`FALSE`
* _categorical_ values, namely characters.
* Further types are only possible by using transformations.

```{r ps}
ps = ParamSet$new(
  params = list(
    ParamInt$new(id = "z", lower = 1, upper = 3),
    ParamDbl$new(id = "x", lower = -10, upper = 10),
    ParamLgl$new(id = "switch"),
    ParamFct$new(id = "methods", values = c("a","b","c"))
  )
)
```

Draw random samples / create random design:

```{r pssample}
smplr = SamplerUnif$new(ps)
smplr$sample(3)

generate_design_random(ps, 3)
```

Generate LHS Design:

```{r pslhs}
generate_design_lhs(ps, 3)
```

Generate Grid Design:

```{r psgrid}
generate_design_grid(ps, resolution = 2)
```

Properties of the parameters within the ParamSet:

```{r psprobs}
ps$ids()
ps$values
ps$nlevels
ps$is_number
```

### Numeric ParamSet

Things you can do on an all numeric ParamSet:

```{r psnumeric}
ps = ParamSet$new(
  params = list(ParamInt$new(id = "z", lower = -10, upper = 10))
)
x_ps = ParamDbl$new(id = "x", lower = 0, upper = 1)$rep(2L)
ps$add(x_ps)
ps$lower
ps$upper
```

The usage of `Parameter$repeat()` generates tags that indicate to which group the parameters belong to:

```{r psnumerictags}
ps$tags
```

This becomes useful if you want to do operations on parameters of one group like with transformations.

### Transformations

Transformations are functions with a fixed signature.

* `x` A `data.table` of parameter values. Each column contains a vector of parameter values of a single parameter.
* `dict` An environment, that can be accessed using the `$` operator. It can contains values that don't belong to any parameter but are important for transformations.
* `tags` A list of the tags for each parameter. Each parameter can have various tags indicating additional characteristics.

Transformations are useful to scale parameters:

```{r pstransscale}
ps = ParamSet$new(
  params = list(
    ParamInt$new(id = "z", lower = -3, upper = 3),
    ParamDbl$new(id = "x", lower = 0, upper = 1)
  )
)
ps$trafo = function(x, param_set) {
  x$z = 2^x$z
  return(x)
}
ps_smplr = SamplerUnif$new(ps)
x = ps_smplr$sample(5)
xt = ps$trafo(x$data)
```

### Advanced Transformations

The following creates a ParamSet with a transformation that scales the `x` values and returns them as a vector.
The original parameters will be removed from the trafo result.
Keep in mind that `z` stays untouched and remains after the transformation.

```{r psnumtrans}
ps = ParamSet$new(
  params = c(
    list(ParamInt$new(id = "z", lower = -10, upper = 10)),
    ParamDbl$new(id = "x", lower = 0, upper = 1)$rep(2L)
  ),
  trafo = function(x, dict, tags) {
    scale = function(x1, x2) c(x1, x2) / sqrt(x1^2+x2^2)
    x$x = Map(scale, x$x_repeated_1, x$x_repeated_2)
    x$x_repeated_1 = NULL
    x$x_repeated_2 = NULL
    return(x)
  }
)
```

The output of all value generating functions won't change for a ParamSet that has a `trafo` function.
Instead these outputs can be put into `ps$transform()` to obtain the desired parameter values.

```{r psnumtrans2}
x = ps$generate_lhs_design(3)
ps$transform(x)
```

For more advanced transformations on repeated parameters you can use `trafo_on_repeated_param()`:

```{r trafo_on_repeated_param}
ps = ParamSet$new(
  params = c(
    list(
      ParamLgl$new(id = "switch"),
      ParamInt$new(id = "z", lower = 1, upper = 4)
    ),
    ParamDbl$new(id = "x", lower = 0, upper = 1)$rep(4L)
  ),
  trafo = trafo_on_repeated_param(
    fun = function(x, dict, tags) {
      scale = function(z, ...) {
        x = c(...)[1:z]
        x / sum(x)
      }
      res = do.call(Map, c(list(f = scale, z = dict$z), as.list(x)))
      list(x = res)
    }, repeated_param_id = "x", additional_params = "z")
  )
(x = ps$sample(3))
ps$transform(x)
```
