---
title: "Tree structured ParamSets"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
    dev: svg
vignette: >
  %\VignetteIndexEntry{Tree structured ParamSets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, cache = FALSE}
library(phng)
set.seed(123)
knitr::opts_chunk$set(cache = TRUE, collapse = FALSE, dev = "svg", fig.height = 3.5)
knitr::knit_hooks$set(document = function(x){
  gsub("```\n*```r*\n*", "", x)
})
```

# ParamTree 

## Support Vector Machine Hyper-parameter Specification
Some hyper-parameters could depend on other hyper-parameters. For example, if we have a model multiplexer of Support Vector Machine(SVM) and Random Forest(RF). The hyper-parameter "kernel" only makes sense if the chosen model is "SVM". Moreover, the hyper-parameter of "gamma" only makes sense if the chosen kernel for "SVM" is "rbf". This kind of relationship resembles a tree or directed graph structure where a node is only activated when its corresponding parent node is activated. 

To tackle the above mentioned problem, we have the following code. Each separate hyper-parameter has a unique identifier called "id" which is defined inside the ParamSimple(ParamCategorical, ParamReal, ..etc). These "id"s are extracted from ParamSimple and used as the unique identifier as the node in the tree as well.

```{r example1}
   library(phng)
   pst = ParamTreeFac(
       ParamCategorical$new(id = "model", values = c("SVM", "RF")),
       addDep(ParamReal$new(id = "C", lower = 0, upper = 100), 
         did = "model", expr = quote(model == "SVM")), # did here means dependant id
       addDep(ParamCategorical$new(id = "kernel", values = c("rbf", "poly")), 
         did = "model", expr = quote(model == "SVM")),
       addDep(ParamReal$new(id = "gamma", lower = 0, upper = 101), 
         did = "kernel", expr = quote(kernel == "rbf")),
       addDep(ParamInt$new(id = "n", lower = 1L, upper = 10L), 
         did = "kernel", expr = quote(kernel == "poly")),
       addDep(ParamInt$new(id = "n_tree", lower = 1L, upper = 10L), 
         did = "model", expr = quote(model == "RF"))
       )
   pst$sample()
   pst$toStringVal()
```

## Neural network architecture specification
```{r}
  ps = recursiveParaFac(2,
      ParamCategorical$new(id = "activation_fun", values = c("sigmoid", "tanh")),
      ParamReal$new(id = "regulation_lambda", lower = 0, upper = 100.0),
      ParamInt$new(id = "n_neurons", lower = 1L, upper = 10L)
      )
  ps$sample()
  ps$toStringVal()
```

## Advanced features

We also provide the user with a "Functional API" which means the user could program at runtime a dynamic structure of the tree. 

Here we introduce the concept of mandatory child node and conditional child node. Mandatory child node means a dependent childnode that is unconditional and conditional child node means the node only make sense once the parent node meet some requirements. For example, in a model multiplexer of "Support Vector Machine" (SVM) and "Random Forest" (RF), the parameter "n_tree" only make sense if the "Model" is "RF". But if the root node is "RF", then "n_tree" becomes a Mandatory child node for "RF".

Secondly, we introduce the concept of ParamHandle, which takes argument of `node` and `depend`. It is basically representing a node in the tree structure which contains pointers to its mandatory child nodes and conditional child nodes.

```{r example2}
  ps = ParamHandle$new(node = ParamCategorical$new(id = "Model", values = c("SVM", "RF")))
  ps$setRoot(ps)  # set the current node to be the root node
  temp_ntree = ParamHandle$new(node = ParamInt$new(id = "n_tree", lower = 1L, upper = 10L), 
    depend = list(id = "Model", fun = quote(Model == "RF")))
  ntree = ps$addCondChild(temp_ntree)  # the condition is contained in temp_ntree§depend already
  temp = ParamHandle$new(node = ParamReal$new(id = "C", lower = 0, upper = 100), 
    depend = list(id = "Model", fun = quote(Model == "SVM")))
  c = ps$addCondChild(temp)
  temp = ParamHandle$new(node = ParamCategorical$new(id = "kernel", values = c("rbf", "poly")),
    depend = list(id = "Model", fun = quote(Model == "SVM")))
  kernel = ps$addCondChild(temp)
  temp = ParamHandle$new(node = ParamReal$new(id = "gamma", lower = 0, upper = 100), 
    depend = list(id = "kernel", fun = quote(kernel == "rbf")))
  gamma = kernel$addCondChild(temp)
  temp = ParamHandle$new(node = ParamInt$new(id = "n", lower = 1L, upper = 10L), 
    depend = list(id = "kernel", fun = quote(kernel == "poly")))
  poly = kernel$addCondChild(temp)
  ps$sample()
  ps$toStringVal() 
  hh = ps$visitor$toFlat()
  poly$getRoot$sample() # use the root to sample
  ps$toStringVal()
  hh = ps$visitor$toFlat()
```

