---
title: "Tree structured ParamSets"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
    dev: svg
vignette: >
  %\VignetteIndexEntry{Tree structured ParamSets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, cache = FALSE}
library(phng)
set.seed(123)
knitr::opts_chunk$set(cache = TRUE, collapse = FALSE, dev = "svg", fig.height = 3.5)
knitr::knit_hooks$set(document = function(x){
  gsub("```\n*```r*\n*", "", x)
})
```

## Functional API

We also provide the user with a "Functional API" which means the user could program at runtime a dynamic structure of the tree. 

Here we introduce the concept of mandatory child node and conditional child node. Mandatory child node means a dependent childnode that is unconditional and conditional child node means the node only make sense once the parent node meet some requirements. For example, in a model multiplexer of "Support Vector Machine" (SVM) and "Random Forest" (RF), the parameter "n_tree" only make sense if the "Model" is "RF". But if the root node is "RF", then "n_tree" becomes a Mandatory child node for "RF".

Secondly, we introduce the concept of ParamHandle, which takes argument of `node` and `depend`. It is basically representing a node in the tree structure which contains pointers to its mandatory child nodes and conditional child nodes.

```{r example2}
  ps = ParamHandle$new(node = ParamCategorical$new(id = "Model", values = c("SVM", "RF")))
  ps$setRoot(ps)  # set the current node to be the root node
  temp_ntree = ParamHandle$new(node = ParamInt$new(id = "n_tree", lower = 1L, upper = 10L), 
    depend = list(id = "Model", fun = quote(Model == "RF")))
  ntree = ps$addCondChild(temp_ntree)  # the condition is contained in temp_ntree§depend already
  temp = ParamHandle$new(node = ParamReal$new(id = "C", lower = 0, upper = 100), 
    depend = list(id = "Model", fun = quote(Model == "SVM")))
  c = ps$addCondChild(temp)
  temp = ParamHandle$new(node = ParamCategorical$new(id = "kernel", values = c("rbf", "poly")),
    depend = list(id = "Model", fun = quote(Model == "SVM")))
  kernel = ps$addCondChild(temp)
  temp = ParamHandle$new(node = ParamReal$new(id = "gamma", lower = 0, upper = 100), 
    depend = list(id = "kernel", fun = quote(kernel == "rbf")))
  gamma = kernel$addCondChild(temp)
  temp = ParamHandle$new(node = ParamInt$new(id = "n", lower = 1L, upper = 10L), 
    depend = list(id = "kernel", fun = quote(kernel == "poly")))
  poly = kernel$addCondChild(temp)
  ps$render2str()
  ps$visitor$toFlat()
  poly$getRoot$render2str() # use the root to sample
  ps$sample(3)
```

