---
title: "ParamSet Tune Short Forms"
author: "Martin Binder"
date: "8/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", collapse = TRUE)
devtools::load_all("..")
library("mlr3")
library("mlr3pipelines")
library("mlr3learners")
library("mlr3filters")
library("mlr3tuning")


p1 = ppl("branch", list(
  "pca" = po("pca"),
  "nothing" = po("nop")
))
p2 = flt("anova")
p3 = ppl("branch", list(
  "svm" = lrn("classif.svm", id = "svm", kernel = "radial",
    type = "C-classification"),
  "xgb" = lrn("classif.xgboost", id = "xgb"),
  "rf" = lrn("classif.ranger",  id = "rf")
), prefix_branchops = "lrn_")
gr = p1 %>>% p2 %>>% p3
glrn = GraphLearner$new(gr)
options(width = 100)
```

## The Problem

We currently have to write a lot to define a tuning paramset. E.g. our presentation for pipelines tuning tunes a stacked learner, and the slide looks like this:

```{r, eval = FALSE}
ps = ParamSet$new(list(
  ParamFct$new("branch.selection", levels = c("pca", "nothing")),
  ParamDbl$new("anova.filter.frac", lower = 0.1, upper = 1),
  ParamFct$new("lrn_branch.selection", levels = c("svm", "xgb", "rf")),
  ParamInt$new("rf.mtry", lower = 1L, upper = 20L),
  ParamInt$new("xgb.nrounds", lower = 1, upper = 500),
  ParamDbl$new("svm.cost", lower = -12, upper = 4),
  ParamDbl$new("svm.gamma", lower = -12, upper = -1)))
ps$add_dep("rf.mtry", "lrn_branch.selection", CondEqual$new("rf"))
ps$add_dep("xgb.nrounds", "lrn_branch.selection", CondEqual$new("xgb"))
ps$add_dep("svm.cost", "lrn_branch.selection", CondEqual$new("svm"))
ps$add_dep("svm.gamma", "lrn_branch.selection", CondEqual$new("svm"))
ps$trafo = function(x, param_set) {
  if (x$lrn_branch.selection == "svm") {
    x$svm.cost = 2^x$svm.cost; x$svm.gamma = 2^x$svm.gamma
  }
  return(x)
}
inst = TuningInstanceSingleCrit$new(tsk("sonar"), glrn, rsmp("cv", folds=3),
 msr("classif.ce"), ps, trm("evals", n_evals = 10))
```

A lot of the information here is redundant, because the `GraphLearner` being tuned already knows a lot about the parameters, their ranges, and their relationships. Furthermore, even writing down a new `ParamSet` takes much more typing than it should. In particular, trafos and dependencies are written at different places than the code which it affects. We identify three problems:

1. Generating `ParamSet`s is verbose
2. Trafos and dependencies are not defined "locally"
3. Specifying tuning `ParamSet`s is redundant.

### Teaser:

How nice would it be if we could write the `ParamSet` like the following:
```{r}
pars <- ps(
  branch.selection = p_fct(c("pca", "nothing")),
  anova.filter.frac = p_dbl(.1, 1),
  lrn_branch.selection = p_fct(c("svm", "xgb", "rf")),
  rf.mtry = p_int(1, 20, requires = lrn_branch.selection == "rf"),
  xgb.nrounds = p_int(1, 500, requires = lrn_branch.selection == "xgb"),
  svm.cost = p_dbl(-12, 4, requires = lrn_branch.selection == "svm",
    trafo = function(x) 2^x),
  svm.gamma = p_dbl(-12, -1, requires = lrn_branch.selection == "svm",
    trafo = function(x) 2^x)
)
```

or, making use of the information from the pipeline `ParamSet`:

```{r}
glrn$param_set$values = list(
  branch.selection = tune(),
  anova.filter.frac = tune(.1, 1),
  lrn_branch.selection = tune(),
  rf.mtry = tune(1, 20),
  xgb.nrounds = tune(1, 500),
  xgb.verbose = 0,
  svm.cost = tune(p_dbl(-12, 4, trafo = function(x) 2^x)),
  svm.type = "C-classification",
  svm.kernel = "radial"
)

glrn$param_set$tune_ps
```

## The Solution

I propose that these problems have solutions that are closely linked. A particular, there is one element in common to them: Parameter ranges, or what I am going to call "`Domain`". (This is probably close, in some way, to what set6 is doing?)

### `Domain`

A `Domain` is basically a `Param` without an ID, but with a trafo and with dependencies. It is an auxiliary object that has not much functionality and should only be used within shortform-functions, the user should not do any computation with them.

This is a nice object to have for the combined problems above, because

1. If we have a non-verbose constructor for `Domain`, then building a `ParamSet` from it could also be non-verbose. Granted, we can also get this by just having shortforms for `ParamDbl$new()` etc., but we won't get the following two benefits.
2. We can use `Domain` to specify trafos and dependencies locally. Instead of defining a `ParamSet` and *then* defining its dependencies and then its trafos, we can build a `ParamSet` from `Domain`s, where each domain contains the trafo and dependency that regards only itself.
3. We can use `Domain` to specify tuning ranges. This is why domains should be unnamed, so they can also be used for specifying tuning ranges. Here we have the benefit that we use a `Domain` object for both quickly defining `ParamSet`s and also for quickly defining tuning ranges.

We get the `Domain` constructors `p_int()`, `p_dbl()`, `p_fct()`, `p_lgl()`, `p_uty()`. We call them just like we call `ParamInt$new()`, *except* without an `id`, and *with* an optional `trafo` and `requirements` argument.

```{r}
single_digits <- p_int(0, 9)
logscale <- p_dbl(log(.01), log(10), trafo = exp)
fac <- p_fct(c("polynomial", "radial"))

# There is some implicit behaviour in that the `p_fct` Domain
# automatically generates a transformation for non-character elements.
# For example, the following:
funfac <- p_fct(c("identity", "log", "exp"),
  trafo = function(x) switch(x,
    identity = identity,
    log = log,
    exp = exp
))
# is much shorter like this:
funfac <- p_fct(list(identity = identity, log = log, exp = exp))

# we can specify requirements for a Domain. Here we say that, whatever
# parameter we define with it, will depend on some "kernel" parameter
# being equal to `"polynomial"`.
degree <- p_int(1, 4,
  requires = kernel == "polynomial" && kernel2 == "polynomial")
```

#### Internals

Because people are not supposed to use `Domain` outside of "sugary" usage, and in particular because they should not do any computation on these objects besides "sugar", we don't need to give these objects much inner life. They are just a `list()`, maybe with a printer (I will avoid calling things like this "`S3` objects" in this document for political reasons), with elements `constructor`, `constargs`, `trafo`, and `requirements`.

Constructing a `Param` from this is just `mlr3misc::invoke(constructor, id = <ID>, .args = constargs)`; the `trafo` and `requirements` will have to be handled in a way to be given to the resulting `ParamSet`.

### `ParamSet` Short Form Construction

We get a function `ps()` that collects `Domain` objects to a complete `ParamSet`. Its arguments must be named. This is very natural and similar to how we would write a list with named arguments etc.

```{r}
pars <- ps(
  a = p_int(1, 10),
  kernel = fac,
  kernel2 = fac,
  kernel3 = fac,
  c = funfac,
  degree = degree
)
pars
```

This is mostly implemented, see how trafo is already working:
```{r}
set.seed(1)
generate_design_random(pars, 1)$transpose()
```

Dependencies also work
```{r}
pars$deps
```

If a `trafo` is needed that goes beyond modifying single parameters, it can be given to an `.extra_trafo` argument to `ps()`. It gets executed *after* the `Domain`-local trafos.

```{r}
pars <- ps(
  a = p_dbl(0, 1, trafo = exp),
  b = p_dbl(0, 1, trafo = exp),
  .extra_trafo = function(x, ps) {
    x$c <- x$a + x$b
    x
  }
)

# See how the addition happens after exp()ing:
pars$trafo(list(a = 0, b = 0))
```

#### Internals

Just call the `Param` constructors from the `Domain` objects as described above. The `trafo` functions of each `Domain` are put together into one big `trafo` for the resulting `ParamSet`, and the dependencies are parsed and added.

### Tune `ParamSet` autogeneration

When defining tuning `ParamSet`s, we want to make use of the information stored in the `Learner`'s `ParamSet`. A nice way to define a tuning scenario is if we can define the fixed and variable parameters of an object at the same time. We solve this by using a `TuneToken` list-with-a-printer, constructed via `tune()`. It is given to the `$values` slot of a `ParamSet` and indicates that a parameter does not have a preset value, and instead should be tuned over.


```{r}
ll <- lrn("classif.rpart")
ll$param_set$values = list(
  minsplit = 10,
  cp = tune()
)
```

The `ParamSet` has a `$tune_ps` active binding that creates the `ParamSet` for tuning out of this. Tuner code, like e.g. `AutoTuner` etc., call this and create a tuning paramset automatically:
```{r}
ll$param_set$tune_ps
```

The `Learner`-side of the `ParamSet` must use `get_values()`, which will throw an error if any `TuneToken` objects are present in the values, since it means the `Learner` is being called with a parameter that should actually be tuned.

The printer of the `TuneToken` can indicate that the respective value of a `ParamSet` is to be tuned:
```{r}
print(ll$param_set$values)
```

Nomenclature: We call `ll$param_set$params$cp` the *underlying parameter*, and `ll$param_set$tune_ps$cp` the *tuning parameter*. They could have different names or types if a `$trafo` is involved.

### `TuneToken`

However, maybe we do not want to tune over the full range of `cp`, or maybe we want to tune over integer values of a `ParamDbl`, or we want to tune over a `ParamUty` with a transformation. The `tune()` constructor therefore admits five behaviours:

1. **`tune()`**: Tune over the whole range of a (bounded) `Param`.
2. **`tune(lower, upper)`**: Tune over the (numeric or integer) `Param` with the given bounds.
3. **`tune(value_vector_or_list)`**: Tune over the values in the given vector or list. This is done by creating a `ParamFct` tuning-`Param` with a trafo that converts to the type required by the underlying parameter.
4. **`tune(Domain)`**: Tune over the domain, making use of the given dependencies and trafos if necessary. This is useful if the type over which we tune is different from the underlying parameter being tuned. See notes below.
5. **`tune(ParamSet)`**: Tune over the `ParamSet`, making use of its `trafo` etc. This is useful if we tune a single (usually `ParamUty`) underlying parameter with multiple tuning parameters.

**Notes**: Why is it nice to have `tune(Domain)` instead of `tune(lower, upper, trafo)` when we need a trafo? Because

1. This functionality is overlapping a lot with `Domain` already, we get two functionalities for the price of one
2. When we give a `trafo`, we can often expect that the tuning parameter and the underlying parameter have different types, e.g. tuning from `log(100)` to `log(1000)` with trafo `round(exp(x))` for a `ParamInt` underlying parameter where the tuning parameter is a `ParamDbl`.

The following performs tuning over the vector-valued `ParamUty` `regularization.factor`; the vector value is constructed from the `reg.sepal` and `reg.petal` tuning parameters.

```{r}
round_exp = function(x) round(exp(x))  # maybe we want this in paradox

lr <- lrn("classif.ranger")
lr$param_set$values = list(
  mtry = 2,
  num.trees = tune(p_dbl(log(10), log(1000), trafo = round_exp)),
  regularization.factor = tune(
    ps(
      reg.sepal = p_dbl(0, 1),
      reg.petal = p_dbl(0, 1),
      .extra_trafo = function(x, param_set) {
        list(regularization.factor =
          c(x$reg.sepal, x$reg.sepal, x$reg.petal, x$reg.petal))
      }
    )
  )
)
lr$param_set$tune_ps

generate_design_random(lr$param_set$tune_ps, 1)$transpose()
```

`mlr3pipelines` has the `affect_columns` parameter, which is a `ParamUty` that takes any object (though often a `Selector` object). Suppose we want to do `PCA` on all columns except one of the `iris` columns:

```{r}
ts = tsk("iris")
glrn = as_learner(po("pca") %>>% lrn("classif.rpart"))
glrn$param_set$values$pca.affect_columns = tune(
  p_fct(ts$feature_names, trafo = function(x) selector_invert(selector_name(x)))
)

generate_design_random(glrn$param_set$tune_ps, 1)$transpose()
```

#### Internals

The `ParamSet$values` slot stores the `TuneToken` and uses that to create a tuning `ParamSet` whenever `$tune_ps` is queried. This is done by
1. Generating a `ParamSet` for each individual value that is set to a `TuneToken`, using the information retrieved from the `TuneToken` (range, factor levels, etc.) and information from the `Param` itself to create the tuning parameter. E.g. `tune()` just clones the `Param`, while `tune(p_fct(...))` needs only the `$id` of the `Param` and does some validity checking.
2. Putting the individual `ParamSets` gotten like this into a common `ParamSet` using a `ps_union()` function. This function goes beyond just collecting the `Param`s, and also collects `$deps` and `$trafo` so that the individual trafos of constituent `ParamSet`s are called correctly. Dependencies of the outer paramset are copied to `$tune_ps`.
3. If we are dealing with a `GraphLearner`, then we are are already dealing with a `ParamSetCollection` on the outside (i.e. `glrn$param_set` is a `ParamSetCollection`). It must provide a `$tune_ps` active binding just as `ParamSet`. It just puts together the individual tuning paramsets using `ps_union()` as well; trafos etc. get handled transparently.

## Documentation

This is how the new features could be documented.

### `Domain`

A `Domain` object is a representation of a single dimension of a `ParamSet`. `Domain` objects are used to construct `ParamSet`s, either through the `ps()` short form, or through the `ParamSet$tune_ps` mechanism. `Domain` corresponds to a `Param` object, except it does not have an `id`, but it *does* have a `trafo` and it does have dependencies (`requires`). For each of the base `Param` classes (`ParamInt`, `ParamDbl`, `ParamLgl`, `ParamFct`, and `ParamUty`) there is a function constructing a `Domain` object (`p_int`, `p_dbl`, `p_lgl`, `p_fct`, `p_uty`). They each have the same arguments as the corresponding `Param` `$new()` function, except without the `id` argument, and with the following additional parameters:

* **`trafo` :: `function`**. Single argument function performing the transformation of a parameter. When the `Domain` is used to construct a `ParamSet`, this transformation will be applied to the corresponding parameter as part of the `$trafo` function.
* **`requires` :: `call`**. An expression indicating a requirement for the parameter that will be constructed from this. Can be given as an expression (using `quote()`), or the expression can be entered directly and will be parsed using NSE (see examples). The expression may be of the form `<Param> == <value>` or `<Param> %in% <values>`, which will result in dependencies according to `ParamSet$add_dep(on = "<Param>", cond = CondEqual$new(<value>))` or `ParamSet$add_dep(on = "<Param>", cond = CondAnyOf$new(<values>))`, respectively. The expression may also contain multiple conditions separated by `&&`.

The `p_fct` function admits a `levels` argument that goes beyond the `levels` accepted by `Paramfct$new()`. Instead of a `character` vector, any atomic vector or list, optionally named, may be given. (If the value is not named, the names are inferred using `as.character()` on the values.) The resulting `Domain` will correspond to a range of values given by the names of the `levels` argument with a `trafo` that maps the `character` names to the arbitrary values of the `levels` argument.

Domain objects are representations of parameter ranges that are intermediate objects to be used in short form constructions in `tune()` and `ps()`. Because of their nature, they should not be modified by the user.

### `ps()`

The `ps()` short form constructor uses `Domain` objects to construct `ParamSet`s in a succinct and readable way. The arguments are:

* **`...` :: `Domain` | `Param`**. Named arguments of `Domain` or `Param` objects. The `ParamSet` will be constructed of the given `Param`s, or of `Params`s constructed from the given domains. The names of the arguments will be used as `id` (the `id` of `Param` arguments are ignored).
* **`.extra.trafo` :: `function(x, ps)`**. Transformation to set the resulting `ParamSet`'s `$trafo` value to. This is in addition to any `trafo` of `Domain` objects given in `...`, and will be run *after* transformations of individual parameters were performed.

### `TuneToken` / `$tune_ps`

A `TuneToken` object can be given to a `ParamSet$values` slot as an alternative to a concrete value. This indicates that the value is not given directly but should be tuned using `mlr3tuning`. If the thus parameterized object is invoked directly, without being wrapped by or given to a tuner, it will give an error.

The tuning range `ParamSet` that is constructed from the `TuneToken` values in a `ParamSet`'s `$values` slot can be accessed through the `ParamSet$tune_ps` active bindng. This is done automatically by tuners if no tuning range is given, but it is also possible to access the `$tune_ps` active binding, modify it further, and give the modified `ParamSet` to a tuning function (or do anything else with it, noone is judging you).

A `TuneToken` represents the range over which the parameter whose `$values` slot it occupies should be tuned over. It can be constructed via the `tune()` function in one of several ways:

* **`tune()`**: Indicates a parameter should be tuned over its entire range. Only applies to finite parameters (i.e. discrete or bounded numeric parameters)
* **`tune(lower, upper)`**: Indicates a numeric parameter should be tuned in the inclusive interval spanning `lower` to `upper`. Depending on the parameter, integer (if it is a `ParamInt`) or real values (if it is a `ParamDbl`) are used.
* **`tune(levels)`**: Indicates a parameter should be tuned through the given discrete values. `levels` can be any named or unnamed atomic vector or list (although in the unnamed case it must be possible to construct a corresponding `character` vector with distinct values using `as.character`). 
* **`tune(<Domain>)`**: The given `Domain` object indicates the range which should be tuned over. The supplied `trafo` function is used for parameter transformation.
* **`tune(<Param>)`**: The given `Param` object indicates the range which should be tuned over.
* **`tune(<ParamSet>)`**: The given `ParamSet` is used to tune over a single `Param`. This is useful for cases where a single evaluation-time parameter value (e.g. `ParamUty`) is constructed from multiple tuner-visible parameters (which may not be `ParamUty`). The supplied `ParamSet` should always contain a `$trafo` function, which must always return a named `list` with a single entry with the name of the `Param` that this `TuneToken` object corresponds to.

## Challenges

* Currently [dependencies on the `Learner`-side are broken](https://github.com/mlr-org/paradox/issues/216), but it should be investigated how they should be handled if we ever get them to work. They are currently added to the result of `$tune_ps` automatically. A question that remains is how dependencies between `Params` that are in different `ParamSets` in a `ParamSetCollection` should be handled, e.g. when a `PipeOp`'s parameter depends on a `PipeOpBranch`. `ParamSetCollection` currently doesn't support this.

* I am not sure whether `trafo = exp` should be allowed for `p_int`, and rounding should happen automatically. An alternative is to create the `round_exp` function as above.
